---
title: "Buzz model"
output: github_document
---

# Markdown version of Buzz data analysis

by: Nina Zumel and John Mount
Win-Vector LLC


To run this example you need a system with R installed (see [http://cran.r-project.org](http://cran.r-project.org)), and data from [https://github.com/WinVector/PDSwR2/tree/master/Buzz](https://github.com/WinVector/PDSwR2/tree/master/Buzz).


We are not performing any new analysis here, just supplying a direct application of Random Forests on the data.

Data from: [http://ama.liglab.fr/datasets/buzz/](http://ama.liglab.fr/datasets/buzz/)
Using: 
       [TomsHardware-Relative-Sigma-500.data](http://ama.liglab.fr/datasets/buzz/classification/TomsHardware/Relative_labeling/sigma=500/TomsHardware-Relative-Sigma-500.data)

(described in [TomsHardware-Relative-Sigma-500.names](http://ama.liglab.fr/datasets/buzz/classification/TomsHardware/Relative_labeling/sigma=500/TomsHardware-Relative-Sigma-500.names) )

Crypto hashes:
shasum TomsHardware-*.txt
*  5a1cc7863a9da8d6e8380e1446f25eec2032bd91  TomsHardware-Absolute-Sigma-500.data.txt
*  86f2c0f4fba4fb42fe4ee45b48078ab51dba227e  TomsHardware-Absolute-Sigma-500.names.txt
*  c239182c786baf678b55f559b3d0223da91e869c  TomsHardware-Relative-Sigma-500.data.txt
*  ec890723f91ae1dc87371e32943517bcfcd9e16a  TomsHardware-Relative-Sigma-500.names.txt

To run this example you need a system with R installed 
(see [cran](http://cran.r-project.org)),
Latex (see [tug](http://tug.org)) and data from 
[PDSwR2](https://github.com/WinVector/PDSwR2/tree/master/Buzz).

To run this example:
* Download buzzm.Rmd and TomsHardware-Relative-Sigma-500.data.txt from the github URL.
* Start a copy of R, use setwd() to move to the directory you have stored the files.
* Make sure knitr is loaded into R ( install.packages('knitr') and
library(knitr) ).
* In R run: (produces [buzzm.md](https://github.com/WinVector/PDSwR2/blob/master/Buzz/buzzm.md) from buzzm.Rmd).
```{r knitsteps,tidy=F,eval=F}
knit('buzzm.Rmd')
```



Now you can run the following data prep steps:

```{r dataprep}
infile <- "TomsHardware-Relative-Sigma-500.data.txt"
paste('checked at', date())
system(paste('shasum', infile), intern=T)  # write down file hash
buzzdata <- read.table(infile, header = FALSE, sep = ",")

makevars <- function(colname, ndays = 7) {
  sprintf("%s_%02g", colname, 0:ndays)
}

varnames <- c(
  "num.new.disc",
  "burstiness",
  "number.total.disc",
  "auth.increase",
  "atomic.containers", # not documented
  "num.displays", # number of times topic displayed to user (measure of interest)
  "contribution.sparseness", # not documented
  "avg.auths.per.disc",
  "num.authors.topic", # total authors on the topic
  "avg.disc.length",
  "attention.level.author",
  "attention.level.contrib"
)

colnames <- unlist(lapply(varnames, FUN=makevars))
colnames <-  c(colnames, "buzz")
colnames(buzzdata) <- colnames

# Split into training and test
set.seed(2362690L)
rgroup <- runif(dim(buzzdata)[1])
buzztrain <- buzzdata[rgroup > 0.1,]
buzztest <- buzzdata[rgroup <=0.1,]
```

This currently returns a training set with `r dim(buzztrain)[[1]]` rows and a test set with 
`r dim(buzztest)[[1]]` rows, which 
`r ifelse(dim(buzztrain)[[1]]==7114 & dim(buzztest)[[1]]==791,'is','is not')` the same
as when this document was prepared.

Notice we have exploded the basic column names into the following:
```{r colnames}
print(colnames)
```

We are now ready to create a simple model predicting "buzz" as function of the
other columns.

```{r model}
# build a model
# let's use all the input variables
nlist = varnames
varslist = as.vector(sapply(nlist, FUN=makevars))

# these were defined previously in Practical Data Science with R
loglikelihood <- function(y, py) {
  pysmooth <- ifelse(py == 0, 1e-12,
                     ifelse(py == 1, 1-1e-12, py))
  sum(y * log(pysmooth) + (1 - y) * log(1 - pysmooth))
}

accuracyMeasures <- function(pred, truth, threshold=0.5, name="model") {
  dev.norm <- -2 * loglikelihood(as.numeric(truth), pred) / length(pred)
  ctable = table(truth = truth,
                 pred = pred > threshold)
  accuracy <- sum(diag(ctable)) / sum(ctable)
  precision <- ctable[2, 2] / sum(ctable[, 2])
  recall <- ctable[2, 2] / sum(ctable[2, ])
  f1 <- 2 * precision * recall / (precision + recall)
  print(paste("precision=", precision, "; recall=" , recall))
  print(ctable)
  data.frame(model = name, 
             accuracy = accuracy, 
             f1 = f1, 
             dev.norm = dev.norm,
             AUC = sigr::calcAUC(pred, truth))
}


library("randomForest")
bzFormula <- paste('as.factor(buzz) ~ ', paste(varslist, collapse = ' + '))
fmodel <- randomForest(as.formula(bzFormula),
                      data = buzztrain,
                      importance = TRUE)

print('training')
rtrain <- data.frame(truth = buzztrain$buzz, 
                     pred = predict(fmodel, newdata = buzztrain, type="prob")[, 2, drop = TRUE])
print(accuracyMeasures(rtrain$pred, rtrain$truth))
WVPlots::ROCPlot(rtrain, "pred", "truth", TRUE, "RF train performance, large model")

print('test')
rtest <- data.frame(truth = buzztest$buzz, 
                    pred = predict(fmodel, newdata=buzztest, type="prob")[, 2, drop = TRUE])
print(accuracyMeasures(rtest$pred, rtest$truth))
WVPlots::ROCPlot(rtest, "pred", "truth", TRUE, "RF train performance, large model")
```

Notice the extreme fall-off from training to test performance, the random forest
over fit on training.  We see good accuracy on test (around 92%), but not the 
perfect fit seen on training.

To try and control the over-fitting we build a new model with the tree
complexity limited to 50 nodes and the node size to at least 100.
This is not necessarily a better model (in fact it scores slightly
poorer on test), but it is one where the training procedure didn't
have enough freedom to memorize the training data (and therefore maybe
had better visibility into some trade-offs.

```{r model2}
fmodel <- randomForest(as.formula(bzFormula),
                      data = buzztrain,
                      maxnodes = 50,
                      nodesize = 100,
                      importance = TRUE)

print('training')
rtrain <- data.frame(truth = buzztrain$buzz, 
                     pred = predict(fmodel, newdata=buzztrain, type="prob")[, 2, drop = TRUE])
print(accuracyMeasures(rtrain$pred, rtrain$truth))

print('test')
rtest <- data.frame(truth = buzztest$buzz, 
                    pred = predict(fmodel, newdata=buzztest, type="prob")[, 2, drop = TRUE])
print(accuracyMeasures(rtest$pred, rtest$truth))
```

And we can also make plots.

Training performance:

```{r plottrain}
WVPlots::ROCPlot(rtrain, "pred", "truth", TRUE, "RF train performance, simpler model")
```

Test performance:

```{r plottest,}
WVPlots::ROCPlot(rtest, "pred", "truth", TRUE, "RF test performance, simpler model")
```

Notice this has similar test performance as the first model.  This
is typical of random forests: the degree of over-fit in the training data
is not a good predictor of good or bad performance on test data.

So we are now left with a choice, unfortunately without clear guidance: which model do we use?
In this case we are going to say: use the simpler model fit on all the data.  The idea
is the simpler model didn't test much worse and more data lets helps the model reduce over-fitting.

```{r model3}
# train on all the data
fmodel <- randomForest(as.formula(bzFormula),
                      data = buzzdata,
                      maxnodes = 50,
                      nodesize = 100,
                      importance = T)
```




Save a prepared R environment.

```{r save,message=F,eval=T}
fname <- 'thRS500.Rdata'
items <- c("varslist", "fmodel", "buzztrain")
save(list = items, file = fname)
message(paste('saved', fname))  # message to running R console
print(paste('saved', fname))    # print to document

paste('finished at', date())
system(paste('shasum', fname), intern = TRUE)  # write down file hash
```

