
# Markdown version of Buzz data analysis

This is the markdown version of the buzz example.  The primary
example is latex ( [pdf result](https://github.com/WinVector/zmPDSwR/blob/master/Buzz/buzz.pdf) ).

We are not perfoming any new analysis here, just supplying a direct application of Random Forests on the data.

Data from: [http://ama.liglab.fr/datasets/buzz/](http://ama.liglab.fr/datasets/buzz/)
Using: 
       [TomsHardware-Relative-Sigma-500.data](http://ama.liglab.fr/datasets/buzz/classification/TomsHardware/Relative_labeling/sigma=500/TomsHardware-Relative-Sigma-500.data)

(described in [TomsHardware-Relative-Sigma-500.names](http://ama.liglab.fr/datasets/buzz/classification/TomsHardware/Relative_labeling/sigma=500/TomsHardware-Relative-Sigma-500.names) )

Crypto hashes:
shasum TomsHardware-*.txt
*  5a1cc7863a9da8d6e8380e1446f25eec2032bd91  TomsHardware-Absolute-Sigma-500.data.txt
*  86f2c0f4fba4fb42fe4ee45b48078ab51dba227e  TomsHardware-Absolute-Sigma-500.names.txt
*  c239182c786baf678b55f559b3d0223da91e869c  TomsHardware-Relative-Sigma-500.data.txt
*  ec890723f91ae1dc87371e32943517bcfcd9e16a  TomsHardware-Relative-Sigma-500.names.txt

To run this example you need a system with R installed 
(see [cran](http://cran.r-project.org)),
Latex (see [tug](http://tug.org)) and data from 
[zmPDSwR](https://github.com/WinVector/zmPDSwR/tree/master/Buzz).

To run this example:
* Download buzzm.Rmd and TomsHardware-Relative-Sigma-500.data.txt from the github URL.
* Start a copy of R, use setwd() to move to the directory you have stored the files.
* Make sure knitr is loaded into R ( install.packages('knitr') and
library(knitr) ).
* In R run: (produces [buzzm.md](https://github.com/WinVector/zmPDSwR/blob/master/Buzz/buzzm.md) from buzzm.Rmd).
```{r knitsteps,tidy=F,eval=F}
knit('buzzm.Rmd')
```


Now you can run the following data prep steps:

```{r dataprep,tidy=F}
buzzdata <- read.table("TomsHardware-Relative-Sigma-500.data.txt", 
                      header=F, sep=",")

makevars <- function(colname, ndays=7) {
  paste(colname, 0:ndays, sep='')
}

varnames <- c("num.new.disc",
             "burstiness",
             "number.total.disc",
             "auth.increase",
             "atomic.containers", # not documented
             "num.displays", # number of times topic displayed to user (measure of interest)
             "contribution.sparseness", # not documented
             "avg.auths.per.disc",
             "num.authors.topic", # total authors on the topic
             "avg.disc.length",
             "attention.level.author",
             "attention.level.contrib"
)

colnames <- as.vector(sapply(varnames, FUN=makevars))
colnames <-  c(colnames, "buzz")
colnames(buzzdata) <- colnames

# Split into training and test
set.seed(2362690L)
rgroup <- runif(dim(buzzdata)[1])
buzztrain <- buzzdata[rgroup > 0.1,]
buzztest <- buzzdata[rgroup <=0.1,]
```

This currently returns a training set with `r dim(buzztrain)[[1]]` rows and a test set with 
`r dim(buzztest)[[1]]` rows, which 
`r ifelse(dim(buzztrain)[[1]]==7114 & dim(buzztest)[[1]]==791,'is','is not')` the same
as when this document was prepared.

Notice we have exploded the basic column names into the following:
```{r colnames,tidy=F}
print(colnames)
```

We are now ready to create a simple model predicting "buzz" as function of the
other columns.

```{r model,tidy=F}
# build a model
# let's use all the input variables
nlist = varnames
varslist = as.vector(sapply(nlist, FUN=makevars))

# these were defined previously,  in Chapter 9
loglikelihood <- function(y, py) {
  pysmooth <- ifelse(py==0, 1e-12,
                     ifelse(py==1, 1-1e-12, py))
  sum(y * log(pysmooth) + (1-y)*log(1 - pysmooth))
}
accuracyMeasures <- function(pred, truth, threshold=0.5, name="model") {
  dev.norm <- -2*loglikelihood(as.numeric(truth), pred)/length(pred)
  ctable = table(truth=truth,
                 pred=pred)
  accuracy <- sum(diag(ctable))/sum(ctable)
  precision <- ctable[2,2]/sum(ctable[,2])
  recall <- ctable[2,2]/sum(ctable[2,])
  f1 <- precision*recall
  print(paste("precision=", precision, "; recall=" , recall))
  print(ctable)
  data.frame(model=name, accuracy=accuracy, f1=f1, dev.norm)
}


library(randomForest)
bzFormula <- paste('as.factor(buzz) ~ ',paste(varslist,collapse=' + '))
fmodel <- randomForest(as.formula(bzFormula),
                      data=buzztrain,
                      mtry=floor(sqrt(length(varslist))),
                      ntree=101,
                      importance=T)

print('training')
rtrain <- data.frame(truth=buzztrain$buzz, pred=predict(fmodel, newdata=buzztrain))
print(accuracyMeasures(rtrain$pred, rtrain$truth))

print('test')
rtest <- data.frame(truth=buzztest$buzz, pred=predict(fmodel, newdata=buzztest))
print(accuracyMeasures(rtest$pred, rtest$truth))
```

Notice the extreme fall-off from training to test performance, the random forest
over fit.  In fact the random forest fit all the data if it sees it during training:

```{r modelA,tidy=F}
fmodel <- randomForest(as.formula(bzFormula),
                      data=buzzdata,
                      mtry=floor(sqrt(length(varslist))),
                      ntree=101,
                      importance=T)
print('all data')
rall <- data.frame(truth=buzztrain$buzz, pred=predict(fmodel, newdata=buzztrain))
print(accuracyMeasures(rall$pred, rall$truth))
```

To try and control the over-fitting we build a new model with the tree
complexity limited to 100 nodes and the node size to at least 20.
This is not necessarily a better model (in fact it scores slightly
poorer on test), but it is one where the training procedure didn't
have enough freedom to memorize the training data (and therefore maybe
had visibility into some trade-offs.

```{r model2,tidy=F}
fmodel <- randomForest(as.formula(bzFormula),
                      data=buzztrain,
                      mtry=floor(sqrt(length(varslist))),
                      ntree=101,
                      maxnodes=100,
                      nodesize=20,
                      importance=T)

print('training')
rtrain <- data.frame(truth=buzztrain$buzz, pred=predict(fmodel, newdata=buzztrain))
print(accuracyMeasures(rtrain$pred, rtrain$truth))

print('test')
rtest <- data.frame(truth=buzztest$buzz, pred=predict(fmodel, newdata=buzztest))
print(accuracyMeasures(rtest$pred, rtest$truth))
```

And we can also make plots.

Training performance:
```{r plottrain, tidy=F}
library(ggplot2)
ggplot(rtrain, aes(x=pred, color=(truth==1),linetype=(truth==1))) + 
   geom_density(adjust=0.1,)
```

Test performance:
```{r plottest,tidy=F}
ggplot(rtest, aes(x=pred, color=(truth==1),linetype=(truth==1))) + 
   geom_density(adjust=0.1)
```

Note the classifier scores are concentrated near zero and one
(meaning the printed confusion matrices pretty much capture the whole
story and the density plots or any sort of ROC plot doesn't add much
value in this case).


